---
title: "callie_test"
author: "Callie Steffen,"
date: "3/10/2020"
output: html_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(janitor)
library(tidyverse)
```

```{r, fig.width= 40, fig.height= 20}
# Data Exploration
#drop NAS only want ships with IHS data
#sort by NOAA score reported-Top performers A+- export as csv!
#distribution from all IHS data and breakdown of letter grade by type
#Biggest discrepancies.. biggest players that mis represent?
#Re categorize for smaller ship type samples

# Arrange Ship Stats By Dropping NAs and rename column of compliance. Arrange in descending order so we can get the top performers that have 100

ship_stats <- read_csv("ship_stats.csv")
#ship_stats <- ship_stats %>% 
  #rename(c("com_score_reported" ="compliance score (reported speed)"))

stats_ihs <-ship_stats %>% 
  drop_na() %>% 
  arrange(-com_score_reported)

#111 vessels with A+ /100 score

top_performers <- stats_ihs %>% 
  head(111) %>% #exported as CSV
  group_by(ship_type_ihs) %>% 
  count(ship_type_ihs)

#Quantity of Each letter grade and ship type 

grades <- stats_ihs %>% 
  group_by(grade) %>% 
 count(grade)
 
grade_quantity <- stats_ihs %>% 
  group_by(grade) %>% 
  count(ship_type_ihs)

 #Distrubution of all letter grades by vessel type (how many of each vessel type is getting each letter grade)
letter_plot <- ggplot(grade_quantity, aes(fill= ship_type_ihs, y= n, x=grade)) + 
    geom_bar(position="stack", stat="identity")

letter_plot

  


```

#Interested in seeing distrubution of categories 

```{r}

ihs_type_category <- stats_ihs %>%
  mutate(
    category = case_when(
      ship_type_ihs %in% c("Tug","Articulated Pusher Tug","Pusher Tug","Anchor Handling Tug Supply") ~ "Tug",
      ship_type_ihs %in% c("Bulk Carrier","Vehicles Carrier","Wood Chips Carrier","Heavy Load Carrier","Bulk Carrier, Self-discharging") ~ "Carrier",
      ship_type_ihs %in% c("Crude Oil Tanker","Chemical/Products Tanker", "Crude/Oil Products Tanker", "Products Tanker", "Replenishment Tanker","LPG Tanker","LNG Tanker","Chemical Tanker") ~ "Tanker",
      ship_type_ihs %in% c("General Cargo Ship","Replenishment Dry Cargo Vessel","Open Hatch Cargo Ship","Logistics Vessel (Naval Ro-Ro Cargo)","General Cargo Ship (with Ro-Ro facility)","Refrigerated Cargo Ship", "Ro-Ro Cargo Ship" ) ~ "Cargo",
      ship_type_ihs%in%c("Research Survey Vessel","Fishery Research Vessel") ~"Research",
      ship_type_ihs%in%c( "Yacht","Yacht (Sailing)") ~"Yacht",
      ship_type_ihs%in%c( "Passenger/Cruise","Passenger Ship") ~"Passenger",
      ship_type_ihs%in%c("Landing Craft","Frigate","Work/Maintenance Pontoon,non propelled","Icebreaker","Crew/Supply Vessel","Drilling Ship","Pollution Control Vessel","Cable Layer", "Hopper,Motor","Salvage Ship","Cable Repair Ship","Training Ship","Utility Vessel","Trailing Suction Hopper Dredger") ~ "Other",
      ship_type_ihs%in%c("Platform Supply Ship") ~"Platform Supply Ship",
      ship_type_ihs%in%c("Container Ship (Fully Cellular)","Container Ship (Fully Cellular/Ro-Ro Facility)") ~"Container Ship",
      ship_type_ihs%in%c("Fishing Vessel") ~ "Fishing Vessel"))
   









# Count Grades Per Category
category_quantity <- ihs_type_category %>% 
  group_by(grade) %>% 
  count(category)

# Now Plot For Categories

letter_plot_category <-ggplot(category_quantity, aes(fill= category, y= n, x=grade)) + 
    geom_bar(position="stack", stat="identity")

letter_plot_category

                                       

```

# Categories With Time Discrepancies 

```{r}
reporting_differences <- ihs_type_category %>% 
  mutate(diff_in_score = as.numeric(com_score_reported - `compliance score (calculated speed)`))

category_differnces <- reporting_differences %>% 
  group_by(category) %>%
  summarize(total_cat = sum(diff_in_score))


  

```


#Categories of Main Interst: Tankers, Bulk, and Container - what is all considered "passenger" - Lets look at carrier (bulk), cargo,tankers,container

```{r}
# filter out for categories of interest

interest <- ihs_type_category %>% 
  filter(category == "Cargo"| category == "Carrier" | category == "Tanker" | category == "Container Ship")

```

# How many ship types does each operators have and how many of each?
```{r}
#467 operators total
#530 operators = 63 operators that have more than one ship type

# amount of each ship type for each operator
operators_shiptypes <- stats_ihs %>%
  group_by(operator) %>% 
  count(ship_type_ihs)

```

```{r}

#operator and how many ship types they own
distinct_ships <- stats_ihs %>%
  select(operator,ship_type_ihs,`total distance (nautcal miles)`) %>% 
  group_by(operator) %>% 
  summarise(types = n_distinct(ship_type_ihs))

# total amount of nautical miles travelled by each operator
                               
 distinct_ships_2 <- stats_ihs %>%
  select(operator, `total distance (nautcal miles)`) %>% 
  group_by(operator) %>% 
  summarize(total_trav = sum(`total distance (nautcal miles)`))
 
 ## combine by operator so we can see how many ship types and distance.

operators_shiptype_distance <- full_join(distinct_ships,distinct_ships_2)

#write to csvs

write.csv(operators_shiptype_distance, "operator_shiptype_distance.csv")

write.csv(operators_shiptypes, "operator_type_quantity.csv")

```

#Looking at Big Query Data
```{r, fig.width= 40, fig.height= 20}
big_query <- read_csv("mmsi_cooperation_stats.csv")

#Quantity of Each letter grade and ship type 

grades_BQ <- big_query %>% 
  group_by(grade) %>% 
 count(grade)
 
shiptype_bq <- big_query %>% 
  group_by(grade) %>% 
  count(shiptype)

 #Distrubution of all letter grades by vessel type (how many of each vessel type is getting each letter grade)
letter_plot_bq <- ggplot(shiptype_bq, aes(fill= shiptype, y= n, x=grade)) + 
    geom_bar(position="stack", stat="identity")

letter_plot_bq



  
```

```{r}
shiptype_category_BQ <- big_query %>%
  mutate(
    category = case_when(
      shiptype %in% c("Tug","Articulated Pusher Tug","Pusher Tug","Anchor Handling Tug Supply") ~ "Tug",
      shiptype %in% c("Bulk Carrier","Vehicles Carrier","Wood Chips Carrier","Heavy Load Carrier","Bulk Carrier, Self-discharging") ~ "Carrier",
      shiptype%in% c("Crude Oil Tanker","Chemical/Products Tanker", "Crude/Oil Products Tanker", "Products Tanker", "Replenishment Tanker","LPG Tanker","LNG Tanker","Chemical Tanker") ~ "Tanker",
      shiptype %in% c("General Cargo Ship","Replenishment Dry Cargo Vessel","Open Hatch Cargo Ship","Logistics Vessel (Naval Ro-Ro Cargo)","General Cargo Ship (with Ro-Ro facility)","Refrigerated Cargo Ship", "Ro-Ro Cargo Ship" ) ~ "Cargo",
      shiptype%in%c("Research Survey Vessel","Fishery Research Vessel") ~"Research",
      shiptype%in%c( "Yacht","Yacht (Sailing)") ~"Yacht",
      shiptype%in%c( "Passenger/Cruise","Passenger Ship") ~"Passenger",
      shiptype%in%c("Landing Craft","Frigate","Work/Maintenance Pontoon,non propelled","Icebreaker","Crew/Supply Vessel","Drilling Ship","Pollution Control Vessel","Cable Layer", "Hopper,Motor","Salvage Ship","Cable Repair Ship","Training Ship","Utility Vessel","Trailing Suction Hopper Dredger") ~ "Other",
      shiptype%in%c("Platform Supply Ship") ~"Platform Supply Ship",
      shiptype%in%c("Container Ship (Fully Cellular)","Container Ship (Fully Cellular/Ro-Ro Facility)") ~"Container Ship",
      shiptype%in%c("Fishing Vessel") ~ "Fishing Vessel"))



# Count Grades Per Category
category_quantity_bq <- shiptype_category_BQ %>% 
  group_by(grade) %>% 
  count(category)

# Now Plot For Categories

letter_plot_category_bq <-ggplot(category_quantity_bq, aes(fill= category, y= n, x=grade)) + 
    geom_bar(position="stack", stat="identity")

letter_plot_category_bq

   
```

```{r}
# How many ship types does each operators have and how many of each?

#498 operators total
#573 operators = 75 ship types distributed between operators

# amount of each ship type for each operator
operators_shiptypes_bq <- big_query %>%
  group_by(operator) %>% 
  count(shiptype)

#operator and how many ship types they own
distinct_ships_bq <- big_query %>%
  select(operator,shiptype) %>% 
  group_by(operator) %>% 
  summarise(types = n_distinct(shiptype))

operators_multiple_shiptype <- distinct_ships_bq %>% 
  arrange(-types) %>% 
  head(57)
#57 operators have more than one ship type

# get a list of each operator, how many each type the operator has
multiple_types <- full_join(operators_multiple_shiptype,operators_shiptypes_bq) %>% head(132)

# total amount of nautical miles travelled by each operator
                               
 bq_travel<- big_query %>%
  select(operator, total_distance_km) %>% 
  group_by(operator) %>% 
  summarize(total_trav = sum(total_distance_km))
 
  bq_travel_2<- big_query %>%
  select(operator, total_distance_km_under_10) %>% 
  group_by(operator) %>% 
  summarize(total_under10 = sum(total_distance_km_under_10))
  
  #combine speed, distance, operator, ship type??



```

# Speed
```{r}
speed_operator <- big_query %>%
  select(operator, shiptype, avg_speed_knots) %>% 
  group_by(operator) %>% 
  summarize(speed_oper = mean(avg_speed_knots))
  
```

#Speed Bins- Sum the amount of nautical miles in speed bracket vs Total amount of miles travelled. Then do it by Container Ship, Cargo,Carrier, Tanker
```{r}

#2017-2019
under10 <- sum(big_query$total_distance_km_under_10)
bet10_12 <- sum(big_query$total_distance_km_btwn_10_12)
bet12_15 <- sum(big_query$total_distance_km_btwn_12_15)
over15 <- sum(big_query$total_distance_km_over_15)
total_km <- sum(big_query$total_distance_km)
sums <- sum(under10,bet10_12,bet12_15,over15)
difference <- total_km - sums #7122.4

per_10 <- under10/total_km*100
per_10_12 <- bet10_12/total_km*100
per12_15 <- bet12_15/total_km*100
per_over15 <- over15/total_km*100

total_per <-per_10 + per_10_12 + per12_15 + per_over15 #99.6

speed_bins <- matrix(c(under10,bet10_12,bet12_15,over15,total_km,per_10,per_10_12,per12_15,per_over15,total_per),ncol=5,byrow=TRUE) 
rownames(speed_bins) <- c("Total km","Percentage")
colnames(speed_bins) <- c("<10","10-12","12-15","over 15","Total")

# Look into interest of Containers, Cargo, Carrier Tanker for 2019

container <- shiptype_category_BQ %>% 
  filter(category == "Container Ship") %>% 
  filter(year == "2019")

carrier <- shiptype_category_BQ %>% 
filter(category == "Carrier") %>% 
  filter(year == "2019")

cargo <- shiptype_category_BQ %>% 
  filter(category == "Cargo") %>% 
  filter(year == "2019")

tanker <- shiptype_category_BQ %>% 
  filter(category == "Tanker") %>% 
  filter(year == "2019")

container_distance<- sum(container$total_distance_km_under_10)
container_total <- sum(container$total_distance_km)
container_compliance_2019 <-container_distance/container_total*100

carrier_distance <- sum(carrier$total_distance_km_under_10)
carrier_total <- sum(carrier$total_distance_km)
carrier_compliance_2019 <-carrier_distance/carrier_total*100


cargo_distance <- sum(cargo$total_distance_km_under_10)
cargo_total <- sum(cargo$total_distance_km)
cargo_compliance_2019 <- cargo_distance/cargo_total*100


tanker_distance <- sum(tanker$total_distance_km_under_10)
tanker_total <- sum(tanker$total_distance_km)
tanker_compliance_2019 <- tanker_distance/tanker_total*100


category_2019 <- matrix(c(container_total,carrier_total,cargo_total,tanker_total,container_compliance_2019,carrier_compliance_2019,cargo_compliance_2019,tanker_compliance_2019),ncol=4,byrow=TRUE) 
rownames(category_2019) <- c("Total km Travelled in VSR","Percentage of Compliance")
colnames(category_2019) <- c("Container","Carrier","Cargo","Tanker")


```


#2019, 2018, and 2017 VSR  Compliance
```{r}

compliance_2019 <- big_query %>% 
  filter(year == "2019")

sum(compliance_2019$total_distance_km_under_10)
sum(compliance_2019$total_distance_km)

484559.6/1148232
#42.2%
  
# 2018
compliance_2018 <- big_query%>% 
  filter(year == "2018")

sum(compliance_2018$total_distance_km_under_10)
sum(compliance_2018$total_distance_km)
100375.2/433805.8
#23.1%

#2017
compliance_2017 <- big_query %>% 
  filter(year == "2017")

sum(compliance_2017$total_distance_km_under_10)
sum(compliance_2017$total_distance_km)
52084.3/248650.6
#20.9%

```

# OPERATORS AND DISTANCE IN 2019
```{r}
operators_distance_2019 <- compliance_2019 %>% 
  group_by(operator) %>% 
  summarize(total_km_2019 = sum(total_distance_km)) %>% 
  arrange(-total_km_2019)


# look at all operators 2017, 2018, 2019
operator_distance_2017_2019 <- big_query %>% 
  group_by(operator) %>% 
  summarize(total_km_trav = sum(total_distance_km)) %>% 
  arrange(-total_km_trav)

# individual ship types to get categories
big_query_ship_types <- unique(big_query$shiptype)
big_query_ship_types <- data.frame(big_query_ship_types)
write.csv(big_query_ship_types, "bigquery_shiptypes.csv")


```
  
#Compare NOAA VSR Stats to Big Query
```{r}
#read in NOAA stats

NOAA_2019 <- read.csv("Subset2019VSRdata_cinms.csv") %>% 
  mutate(cooperation = 100 - pdgt_10) %>% 
mutate(percent_coop = cooperation/100) %>% 
  mutate(miles_cooperation = dist_nm * percent_coop)

# Big query is in km so put into NM-divide the length value by 1.852

bq_2019 <- compliance_2019 %>% 
  mutate(total_nm = total_distance_km/1.852)

#Unique MMSIs
#BQ - 987
unique_bq <- unique(bq_2019$mmsi)
unique_bq <- data.frame(unique_bq) %>% 
  mutate(mmsi = unique_bq)
#NOAA - 2789
unique_mmsi_NOAA <- unique(NOAA_2019$mmsi)
unique_mmsi_NOAA <- data.frame(unique_mmsi) %>% 
  mutate(mmsi = unique_mmsi_NOAA)

#ship names and MMSI numbers

ship_names <- NOAA_2019 %>% 
  select(mmsi,name)

#group by mmsi and look at total nm travelled

mmsi_bq <- bq_2019 %>% 
  group_by(mmsi) %>% 
  summarize(total_nm_2019_bq = sum(total_nm), compliance_bq = total_distance_km_under_10/total_distance_km  * 100)

mmsi_NOAA <- NOAA_2019 %>% 
  group_by(mmsi) %>% 
  summarize(total_nm_2019_noaa = sum(dist_nm), compliance_noaa = sum((miles_cooperation)/sum(total_nm_2019_noaa)) * 100)

# Merge DataFrames to Compare
bq_noaa_comparison <- merge(mmsi_bq,mmsi_NOAA, by = "mmsi")
  
#get ship names
comparison <- merge(bq_noaa_comparison,ship_names,by = "mmsi", all.x = FALSE) %>% 
  unique() %>% 
  mutate(difference_nm = abs(total_nm_2019_bq - total_nm_2019_noaa)) %>% 
  mutate(difference_in_compliance = abs(compliance_bq - compliance_noaa)) 
  
reorder <- comparison[c("name", "mmsi", "total_nm_2019_bq","total_nm_2019_noaa","difference_nm","compliance_bq","compliance_noaa","difference_in_compliance")]

reorder$total_nm_2019_bq <- round(reorder$total_nm_2019_bq ,digit=2)
reorder$total_nm_2019_noaa <- round(reorder$total_nm_2019_noaa ,digit=2)
reorder$difference_nm <- round(reorder$difference_nm ,digit=2)
reorder$compliance_bq <- round(reorder$compliance_bq ,digit=2)
reorder$compliance_noaa <- round(reorder$compliance_noaa ,digit=2)
reorder$difference_in_compliance <- round(reorder$difference_in_compliance ,digit=2)

write.csv("reorder", "bq_noaa_VSR_comparison_2019.csv")

# What ships does NOAA have that BQ doesnt  and ones BQ has that NOAA doesnt
matches <- semi_join(unique_bq,unique_mmsi_NOAA, by = "mmsi")

mmsi_bq_missing_from_NOAA <- anti_join(unique_bq,unique_mmsi_NOAA, by = "mmsi") #missing from NOAA

mmsi_NOAA_missing_from_bq <- anti_join(unique_mmsi_NOAA, unique_bq , by = "mmsi") #missing from BQ




```


```{r}

quality_check <- NOAA_2019 %>% 
  filter(mmsi == "3669999")

sum(quality_check$dist_nm) # 1110.4
sum(quality_check$miles_cooperation) #611.3367

611.3376/1110.4

0.5505562
 



```

