---
title: "Run Through"
author: "Sean Goral"
date: "2/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries, include=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(leaflet)
library(RColorBrewer)
library(maptools)
library(sf)
library(rgeos)
library(dplyr)
library(RPostgreSQL)
library(dbplyr)
library(lubridate)
library(units)
library(data.table)
library(parallel)
library(ggplot2)
library(here)
```

## Source R files
```{r, source}
# Database connection function
source('~/github/whalesafe4r/R/db.R')

# Crawl https://ais.sbarc.org/logs_delimited/ for URLs
source('~/github/whalesafe4r/R/crawlers.R')

# Read ais.txt URL into ais_data df and read ais_data df into ais_segs df
source('~/github/whalesafe4r/R/readers.R')

# 1) Loop through ais.txt URLs to return ais_data
# 2) Create ais_segments from ais_data
# 3) Intersect ais_segments and vsr_zones table in the database to create vsr_segments table in he database
source('~/github/whalesafe4r/R/update_ais.R')

# Utility functions for crawler, date and spatial functions, etc...
source('~/github/whalesafe4r/R/utils.R')

# Create summary statistics for merged vsr_sements/ihs_data df
source('~/github/whalesafe4r/R/seg_stats.R')
```

### Initiate connection
```{r, connection}

# Must be on UCSB IP address & database credentials for connection to work. 
# (add to repo s4w_amazon_rds.yml file)
# (add yml file to your gitignore)

con <- db_connect()
```

### Get last read URL from ais_data table
```{r, last read URL}

#Query database to find the 'last read' url from the ais_data table
last_read_url <- dbGetQuery(con, "SELECT MAX(url) AS last_read_url FROM ais_data;") %>% .$last_read_url

last_read_url

```

### Get list of new links
```{r, new links}
# Use last_read_url in the get_ais_urls() function to get all new ais urls

new_links <- get_ais_urls(last_read_url)

new_links

# Example get_ais_urls()

example_path <- "https://ais.sbarc.org/logs_delimited/2019/191110/AIS_SBARC_191110-00.txt"

example_links <- get_ais_urls(example_path)

example_links %>% head() 
example_links%>% tail()
```

### Get new_ais_data
```{r, new_ais_data}

# Use update_ais_data() to loop through 'new_links' and return new_ais_data data.frame

new_ais_data <- update_ais_data(links = new_links)

example_ais_data <- update_ais_data(links = example_links)

example_ais_data_5000 <- example_ais_data %>% 
  head(5000)

example_ais_data_table <- kable(head(example_ais_data, n = 100L)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")

example_ais_data_table

ex_ais_data_map <- leaflet(example_ais_data) %>% 
        addTiles() %>% 
        addCircles(lat = example_ais_data$lat, lng = example_ais_data$lon) %>% 
        setView(lng = '-119', lat = '34',zoom = 7)

ex_ais_data_map
```

### Append ais_data table in database with new_ais_data
```{r, Append ais_data table}

# Append new_ais_data to ais_data table in the database
# Appending to ais_data1 for testing
dbWriteTable(con = con, 
             name = 'ais_data1', 
             value = new_ais_data, 
             append = TRUE)
```

### Get new_segs_data for global environment
```{r, new_segs_data}

# Use new_ais_data in the update_segments_data() function to return new_segs_data 
new_segs_data <- update_segments_data(con = con, 
                                      ais_data = new_ais_data)

# Use example_ais_data in the update_segments_data() function to return example_segs_data 
example_segs_data <- update_segments_data(con = con, ais_data = example_ais_data_5000)

# Example Map
ex_seg_map <- leaflet(example_segs_data) %>% 
                    setView(lng = '-119', lat = '34',zoom = 7) %>% 
                    addTiles() %>% 
                    addPolylines()

ex_seg_map
```

### Write ais_segs_data to ais_segments table in database
```{r, Write ais_segs_data to database}

# Append new_segs_data to the database if new_segs_data exists
write_new_segs_data <- function(data=new_segs_data){
    if (is.null(new_segs_data)==TRUE){
     print("Null segments data")
    } else {
      dbWriteTable(con = con, 
                   name = 'ais_segments1', 
                   value = new_segs_data, 
                   append = TRUE)
    }
  }

# Run it
write_new_segs_data()
```

### Update vsr_segments table in the database
```{r, Update vsr_segments}

# Update vsr_segments on the database side
update_vsr_segments(con = con)

con <- db_connect()
vsr_segments_ex <- sf::st_read(dsn = con, EWKB = TRUE, 
                               query = "SELECT * FROM vsr_segments 
                                        LIMIT 1000;")

vsr_zones <- sf::st_read(dsn = con, EWKB = TRUE, 
                               query = "SELECT * FROM vsr_zones;")
vsr_2019 <- head(vsr_zones,n = 1L)

ex_vsr_seg_map <- leaflet(vsr_segments_ex) %>% 
  setView(lng = '-119', lat = '34',zoom = 7) %>% 
  addTiles() %>% 
  addPolylines(label = ~sprintf("%0.03f km/hr on %s", seg_kmhr, beg_dt, name), group = "segments") %>% 
  addPolygons(data = vsr_2019, fillColor = 'blue', group = 'vsr') %>%   
  addLayersControl(overlayGroups = c("segments","vsr"))

ex_vsr_seg_map
```

### Merge vsr_segments data with IHS data
```{r, Merge vsr_segments with IHS data}

# Merge vsr_segments with ihs_data from database to return vsr_segs_ihs in global environment. (Takes a little time ~40 seconds)
system.time({
vsr_segs_ihs <- merge_ihs_vsr()
})
```

### Calculate ship statistics
```{r, ship statistics} 

# Calculate ship stats for each mmsi using merged df, vsr_segs_ihs
ship_stats <- ship_statistics(data = vsr_segs_ihs)

ship_stats_2019 <- ship_statistics(data = vsr_segs_ihs,
                                   date_start = '2019-01-01',
                                   date_end = '2019-12-31',
                                   tonnage = 300)

ship_stats_2019_table <- ship_stats_2019 %>% 
  kable(escape = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  scroll_box(width = "100%", height = "300px")

ship_stats_2019_table
```

## Overwrite the ship_stats table in the database 
```{r, ship_stats db write}
# And overwrite the ship_stats table in the database 
dbWriteTable(con = con, 
             name = "ship_stats", 
             value = ship_stats, 
             overwrite = TRUE)
```

### Calculate operator statistics
```{r, operator statistics} 

# Calculate operator stats for each mmsi 
operator_stats = operator_statistics(data = vsr_segs_ihs)

operator_stats_2019 = operator_statistics(data = vsr_segs_ihs,
                                     date_start = '2019-01-01',
                                     date_end = '2019-12-31',
                                     tonnage = 300) %>% filter("total distance (km)" >= 100)

operator_stats_2019_table <- operator_stats_2019 %>% 
  kable(escape = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive")) %>% 
  scroll_box(width = "100%", height = "300px")

operator_stats_2019_table
```

## Overwrite the operator_stats table in the database
```{r, overwrite operator_stats in db}
# And overwrite the operator_stats table in the database 
dbWriteTable(con = con, 
             name = "operator_stats", 
             value = operator_stats, 
             overwrite = TRUE)
```

### Disconnect from database
```{r, disconnect}

dbDisconnect(con)
```

